---
title: "Waterbird extrapolation"
author: "Dana Lanceman"
format: html
editor: visual
---

## Waterbird extrapolation

Aerial waterbird surveys produce estimates of waterbird counts for wetland complexes. However, surveys cannot cover all areas of a large wetland, and survey tracks differ each year, so there is a need to standardise waterbird counts through extrapolation. Here, we demonstrate a method for extrapolating NSW aerial waterbird survey data using the Great Cumbung Swamp as an example. 

This script is run assuming that inundation rasters have already been generated for each year of the Great Cumbung Swamp.

## 1. Set up

First, let's load the packages we need for our analysis.

```{r Packages, echo = FALSE}
library(tidyverse)
library(sf)
library(readxl)
library(raster)
library(terra)
library(here)
```

Next, let's import the datasets we need for this workflow:

-   a shapefile of the Great Cumbung Swamp, with lakes clipped out (cumbung_eco)

    -   note, this says it has 3 features because the process of clipping lakes made some parts disconnected from the rest, but we are treating this as one feature

-   a shapefile of the Great Cumbung Swamp, with lakes and Baconian swamp clipped out (cumbung_noBac), for years where Baconian Swamp was surveyed and should be extrapolated separately

-   a shapefile of Baconian Swamp (Baconian), for years where Baconian Swamp was surveyed and should be extrapolated separately

-   shapefiles with combinations of Baconian swamp, Marool Lake and Blindbungi swamp included as part of the cumbung ecological boundary (cumbung_M, cumbung_B, cumbung_MB, cumbung_Bac_M, cumbung_Bac_B, cumbung_Bac_M) for years where lakes were counted as part of transects, so need to be included in the inundation area

-   an excel file of waterbird records, with named locations, for each year (wb_2010, etc). Note this is not necessary as this step is being done within TDW now. But if you have the data, you can run them here as a check to make sure TDW is processing data the same way.

```{r Import_boundary_waterbirds}
setwd(here("data/"))

cumbung_eco <- st_read("boundary/Cumbung_eco_Blindbungi_dissolved.shp")
cumbung_noBac <- st_read("boundary/Cumbung_eco_noBac.shp")
Baconian <- st_read("boundary/Baconian_Swamp_boundary.shp")
cumbung_noMarool <- st_read("boundary/Cumbung_eco_noMarool.shp")
cumbung_noBac_noM <- st_read("boundary/Cumbung_eco_noBac_noMarool.shp")

wb_2010 <- read_excel("waterbirds/2010_Cumbung_PDF-data.xlsx")
wb_2011 <- read_excel("waterbirds/2011_Cumbung_PDF-data.xlsx")
wb_2016 <- read_excel("waterbirds/2016_Cumbung_PDF-data.xlsx")
```

Import tracklogs.

-   a shapefile of the aerial survey tracklog for each year, pre-clipped in QGIS for actual surveyed tracklog areas (rather than areas where the plane flew, but waterbirds were not surveyed)

```{r Import_tracklogs}
setwd(here("data/tracklogs/Use_these/"))

# 2010 - clarify whether these are meandering or transects. only one count
trk_2010 <- st_read("2010_lines_for_ecological_transects.shp")
trk_2010_AA_C1 <- st_read("2010_Cumbung_AA_C1.shp")
trk_2010_AA_C2 <- st_read("2010_Cumbung_AA_C2.shp")

# 2011 - C1 and C2 combined, but also blindbungi extra with C1 and C2 separate
trk_2011 <- st_read("2011_lines_for_ecological_transects.shp")
trk_2011_blind_C1 <- st_read("2011-Cumbung-tracklog-C1-blindbungi.shp")
trk_2011_blind_C2 <- st_read("2011-Cumbung-tracklog-C2-blindbungi.shp")

# 2012
trk_2012_C1 <- st_read("2012_C1_lines.shp")
trk_2012_C2 <- st_read("2012_C2_lines.shp")

# 2013 - C1 and C2 combined
trk_2013 <- st_read("2013_layers_for_ecological_transects.shp")

# 2014 - C1 and C2 combined
trk_2014 <- st_read("2014_lines_for_ecological_transects.shp")
trk_2014_blind_C1 <- st_read("2014-Cumbung-tracklog-blindbungi-C1-path.shp")
trk_2014_blind_C2 <- st_read("2014-Cumbung-tracklog-blindbungi-C2-path.shp")

# 2015 - C1 and C2 combined
trk_2015 <- st_read("2015_lines_for_ecological_transects.shp")

# 2016 - C1 and C2 are combined in the meandering tracklogs. 20161107 covers Baconian swamp.
trk_2016_transect <- st_read("2016_transects_linestring.shp")
trk_20161109_mean <- st_read("20161109_lines_for_ecological_transects.shp")
trk_20161107_Bac_C1 <- st_read("20161107_BaconianSwamp_C1_lines.shp")
trk_20161107_Bac_C2 <- st_read("20161107_BaconianSwamp_C2_lines.shp")

# 2017 - C1 and C2 combined
trk_2017 <- st_read("2017_lines_for_ecological_transects.shp")
trk_2017_blind_C1 <- st_read("2017_Cumbung_Blindbungi_C1.shp")
trk_2017_blind_C2 <- st_read("2017_Cumbung_Blindbungi_C2.shp")

# 2018 - C1 and C2 combined
trk_2018 <- st_read("2018_lines_for_ecological_transects.shp")

# 2019 - C1 and C2 combined
trk_2019 <- st_read("2019_lines_for_ecological_transects.shp")

# 2020 - C1 and C2 combined
trk_2020 <- st_read("2020_lines_for_ecological_transects.shp")

# 2021 - C1 and C2 combined
trk_2021 <- st_read("2021_lines_for_ecological_transects.shp")

# 2022 - transects
trk_2022 <- st_read("2022_lines_for_ecological_transects.shp")

# 2023 - transects and meandering
trk_2023 <- st_read("2023_transect_paths.shp")

# 2024 - C1 meandering covers more sections than C2
trk_2024_transect <- st_read("2024-cumbung-transect.shp")
trk_2024_C1 <- st_read("2024-cumbung-meandering-C1.shp")
trk_2024_C2 <- st_read("2024-cumbung-meandering-C2.shp")
trk_2024_Baconian <- st_read("2024_Baconian_line.shp")

# 2025 
trk_2025 <- st_read(here("../../2026_Cumbung_Marshes_extrapolation/Data/tracklogs/2025-Cumbung-tracklog-clip_c/2025-Cumbung-tracklog-clip_c.shp"))
```

Import inundation data.

-   an inundation raster for the Great Cumbung Swamp, for each year

    -   note these are for a broader cumbung boundary and will need to be clipped to the cumbung ecological boundary

```{r Import_inun_data}
setwd(here("data/inundation/"))

inun_2010 <- raster("wi_cumbung_20101203_reclass-400000_clip.tif")
inun_2011 <- raster("wi_cumbung_20111104_reclass-550000_clip.tif")
inun_2012 <- raster("wi_cumbung_20121114_reclass-300000_clip.tif")
inun_2013 <- raster("wi_cumbung_20131125_reclass-600000_clip.tif")
inun_2014 <- raster("Cumbung_inundation_20141112_SEED_clip.tif")
inun_2015 <- raster("Cumbung_inundation_20151115_SEED_clip.tif")
inun_2016 <- raster("Cumbung_inundation_20161117_SEED_clip.tif")
inun_2017 <- raster("wi_cumbung_20171104_reclass-670000_clip.tif")
inun_2018 <- raster("Cumbung_inundation_20181022_SEED_clip.tif")
inun_2019 <- raster("Cumbung_inundation_20191112_SEED_clip.tif")
inun_2020 <- raster("Cumbung_inundation_20201002_SEED_clip.tif")
inun_2021 <- raster("Cumbung_inundation_20211022_SEED_clip.tif")
inun_2022 <- raster("wi_cumbung_20221118_reclass-500000_clip.tif")
inun_2023 <- raster("wi_cumbung_20231105_reclass-700000_clip.tif")
inun_2024 <- raster("wi_cumbung_20241030_reclass_-800000_clip.tif")

inun_2025 <- raster(here("../../2026_Cumbung_Marshes_extrapolation/spatial/tiff/cumbung_20251118_reclass_-700000_clip.tif"))
```

## 2. Prepare data

Next, we need to prepare our data.

-   transform everything into the same coordinate system (GDA94 MGA zone 55, which is EPSG 28355)

```{r Transform_coords}
cumbung_eco <- st_transform(cumbung_eco, 28355)  
cumbung_noBac <- st_transform(cumbung_noBac, 28355)
Baconian <- st_transform(Baconian, 28355)
cumbung_noBac_noM <- st_transform(cumbung_noBac_noM, 28355)
cumbung_noMarool <- st_transform(cumbung_noMarool, 28355)

trk_2010 <- st_transform(trk_2010, 28355) 
trk_2010_AA_C1 <- st_transform(trk_2010_AA_C1, 28355)
trk_2010_AA_C2 <- st_transform(trk_2010_AA_C2, 28355)
trk_2011 <- st_transform(trk_2011, 28355)
trk_2011_blind_C1 <- st_transform(trk_2011_blind_C1, 28355)
trk_2011_blind_C2 <- st_transform(trk_2011_blind_C2, 28355)
trk_2012_C1 <- st_transform(trk_2012_C1, 28355)
trk_2012_C2 <- st_transform(trk_2012_C2, 28355)
trk_2013 <- st_transform(trk_2013, 28355)
trk_2014 <- st_transform(trk_2014, 28355)
trk_2014_blind_C1 <- st_transform(trk_2014_blind_C1, 28355)
trk_2014_blind_C2 <- st_transform(trk_2014_blind_C2, 28355)
trk_2015 <- st_transform(trk_2015, 28355)
trk_2016_transect <- st_transform(trk_2016_transect, 28355)
trk_20161107_Bac_C1 <- st_transform(trk_20161107_Bac_C1, 28355) 
trk_20161107_Bac_C2 <- st_transform(trk_20161107_Bac_C2, 28355)
trk_20161109_mean <- st_transform(trk_20161109_mean, 28355)
trk_2017 <- st_transform(trk_2017, 28355)
trk_2017_blind_C1 <- st_transform(trk_2017_blind_C1, 28355)
trk_2017_blind_C2 <- st_transform(trk_2017_blind_C2, 28355)
trk_2018 <- st_transform(trk_2018, 28355)
trk_2019 <- st_transform(trk_2019, 28355)
trk_2020 <- st_transform(trk_2020, 28355)
trk_2021 <- st_transform(trk_2021, 28355)
trk_2022 <- st_transform(trk_2022, 28355)
trk_2023 <- st_transform(trk_2023, 28355)
trk_2024_Baconian <- st_transform(trk_2024_Baconian, 28355)
trk_2024_C1 <- st_transform(trk_2024_C1, 28355)
trk_2024_C2 <- st_transform(trk_2024_C2, 28355)
trk_2024_transect <- st_transform(trk_2024_transect, 28355)
trk_2025 <- st_transform(trk_2025, 28355)

inun_2010 <- projectRaster(inun_2010, crs=28355)
inun_2011 <- projectRaster(inun_2011, crs=28355)
inun_2012 <- projectRaster(inun_2012, crs=28355)
inun_2013 <- projectRaster(inun_2013, crs=28355)
inun_2014 <- projectRaster(inun_2014, crs=28355)
inun_2015 <- projectRaster(inun_2015, crs=28355)
inun_2016 <- projectRaster(inun_2016, crs=28355)
inun_2017 <- projectRaster(inun_2017, crs=28355)
inun_2018 <- projectRaster(inun_2018, crs=28355)
inun_2019 <- projectRaster(inun_2019, crs=28355)
inun_2020 <- projectRaster(inun_2020, crs=28355)
inun_2021 <- projectRaster(inun_2021, crs=28355)
inun_2022 <- projectRaster(inun_2022, crs=28355)
inun_2023 <- projectRaster(inun_2023, crs=28355)
inun_2024 <- projectRaster(inun_2024, crs=28355)
inun_2025 <- projectRaster(inun_2025, crs=28355)

# check coordinate systems are the same
ggplot() + 
  geom_sf(data = cumbung_eco, color = "red") +
  geom_sf(data = trk_2010) +
  geom_sf(data = trk_2010_AA_C1, color = "green") +
  geom_sf(data = trk_2010_AA_C2, color = "blue")

ggplot() + 
  geom_sf(data = cumbung_noMarool, color = "red") +
  geom_sf(data = trk_2011) +
  geom_sf(data = trk_2011_blind_C1, color = "green") +
  geom_sf(data = trk_2011_blind_C2, color = "blue")
  

ggplot() + 
  geom_sf(data = cumbung_noMarool, color = "red") +
  geom_sf(data = trk_2012_C1) +
  geom_sf(data = trk_2012_C2, color = "blue")

ggplot() + 
  geom_sf(data = cumbung_eco, color = "red") +
  geom_sf(data = trk_2013)

ggplot() + 
  geom_sf(data = cumbung_eco, color = "red") +
  geom_sf(data = trk_2014) +
  geom_sf(data = trk_2014_blind_C1, color = "green") +
  geom_sf(data = trk_2014_blind_C2, color = "blue")

ggplot() + 
  geom_sf(data = cumbung_noBac_noM, color = "red") +
  geom_sf(data = trk_2015)
  
ggplot() + 
  geom_sf(data = cumbung_noBac, color = "red") +
  geom_sf(data = trk_2016_transect) +
  geom_sf(data = trk_20161107_Bac_C1, color = "blue") +
  geom_sf(data = trk_20161107_Bac_C2, color = "yellow") +
  geom_sf(data = trk_20161109_mean, color = "green") 

ggplot() + 
  geom_sf(data = cumbung_noBac, color = "red") +
  geom_sf(data = trk_2017) +
  geom_sf(data = trk_2017_blind_C1, color = "blue") +
  geom_sf(data = trk_2011_blind_C2, color = "yellow")

ggplot() + 
  geom_sf(data = cumbung_noBac, color = "red") +
  geom_sf(data = trk_2018)

ggplot() + 
  geom_sf(data = cumbung_eco, color = "red") +
  geom_sf(data = trk_2019)

ggplot() + 
  geom_sf(data = cumbung_noBac, color = "red") +
  geom_sf(data = trk_2020)

ggplot() + 
  geom_sf(data = cumbung_noBac_noM, color = "red") +
  geom_sf(data = trk_2021)

ggplot() + 
  geom_sf(data = cumbung_eco, color = "red") +
  geom_sf(data = trk_2022)

ggplot() + 
  geom_sf(data = cumbung_noBac, color = "red") +
  geom_sf(data = trk_2023)

ggplot() + 
  geom_sf(data = cumbung_noBac_noM, color = "red") +
  geom_sf(data = Baconian, color = "black") +
  geom_sf(data = trk_2024_Baconian, color = "pink") +
  geom_sf(data = trk_2024_C1, color = "blue") +
  geom_sf(data = trk_2024_C2, color = "green") +
  geom_sf(data = trk_2024_transect, color = "yellow") 

ggplot() + 
  geom_sf(data = cumbung_noBac, color = "red") +
  geom_sf(data = trk_2025)
```

-   clip and buffer tracklogs

    -   for cases where they are merged, separate the tracklogs from different dates and/or counts and/or transect/meandering types

    -   buffer distance is 100m for transects and 150m for meandering tracklogs

    -   merge meandering and transect tracklogs where both types have been measured (for C1 and C2 separately)

```{r Prepare_tracklog_data}
# 2010
trk_2010_buff <- trk_2010 %>% 
  st_buffer(dist = 150)
trk_2010_AA_C1_buff <- trk_2010_AA_C1 %>% 
  st_buffer(dist = 150)
trk_2010_AA_C2_buff <- trk_2010_AA_C2 %>% 
  st_buffer(dist = 150)

# bind some buffered tracklogs together (all count 1 and all count 2) for calculations later
trk_2010_C1_all_buff <- bind_rows(trk_2010_AA_C1_buff, trk_2010_buff)
trk_2010_C2_all_buff <- bind_rows(trk_2010_AA_C2_buff, trk_2010_buff)


# 2011
trk_2011_C1_buff <- trk_2011 %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2011_blind_C1_buff <- trk_2011_blind_C1 %>% 
  st_buffer(dist = 150)
trk_2011_C2_buff <- trk_2011 %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)
trk_2011_blind_C2_buff <- trk_2011_blind_C2 %>% 
  st_buffer(dist = 150)

# bind all count 1 and count 2 together
trk_2011_C1_all_buff <- bind_rows(trk_2011_C1_buff, trk_2011_blind_C1_buff)
trk_2011_C2_all_buff <- bind_rows(trk_2011_C2_buff, trk_2011_blind_C2_buff)


# 2012
trk_2012_C1_buff <- trk_2012_C1 %>% 
  st_buffer(dist = 150)

trk_2012_C2_buff <- trk_2012_C2 %>% 
  st_buffer(dist = 150)

# 2013
trk_2013_C1_buff <- trk_2013 %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2013_C2_buff <- trk_2013 %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)

# 2014
trk_2014_C1_buff <- trk_2014 %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2014_C2_buff <- trk_2014 %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 75) # one observer's data is missing, so half the buffer size
trk_2014_blind_C1_buff <- trk_2014_blind_C1 %>% 
  st_buffer(dist = 150)
trk_2014_blind_C2_buff <- trk_2014_blind_C2 %>% 
  st_buffer(dist = 150)

# bind all count 1 and count 2 together
trk_2014_C1_all_buff <- bind_rows(trk_2014_C1_buff, trk_2014_blind_C1_buff)
trk_2014_C2_all_buff <- bind_rows(trk_2014_C2_buff, trk_2014_blind_C2_buff)

# 2015
trk_2015_main_C1_buff <- trk_2015 %>% 
  subset(TransLabel != "Baconian Swamp") %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2015_main_C2_buff <- trk_2015 %>% 
  subset(TransLabel != "Baconian Swamp") %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)
trk_2015_Bac_C1_buff <- trk_2015 %>% 
  subset(TransLabel == "Baconian Swamp") %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2015_Bac_C2_buff <- trk_2015 %>% 
  subset(TransLabel == "Baconian Swamp") %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)

# 2016
trk_2016_transect_buff <- trk_2016_transect %>% 
  st_buffer(dist = 100)

trk_20161107_Bac_C1_buff <- trk_20161107_Bac_C1 %>% 
  st_buffer(dist = 150)

trk_20161107_Bac_C2_buff <- trk_20161107_Bac_C2 %>% 
  st_buffer(dist = 150)

trk_20161109_C1_buff <- trk_20161109_mean %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)

trk_20161109_C2_buff <- trk_20161109_mean %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)

# bind some buffered tracklogs together (all count 1 and all count 2, by date) for calculations later
# transects are part of 20161109
trk_20161109_C1_all <- bind_rows(trk_20161109_C1_buff, trk_2016_transect_buff)
trk_20161109_C2_all <- bind_rows(trk_20161109_C2_buff, trk_2016_transect_buff)
# 20161107 just has Baconian so doesn't need to be further subset/merged


# 2017
trk_2017_Bac_buff <- trk_2017 %>% 
  subset(TransLabel == "Baconian Swamp") %>% 
  st_buffer(dist = 150)
trk_2017_main_C1_buff <- trk_2017 %>% 
  subset(TransLabel != "Baconian Swamp") %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2017_blind_C1_buff <- trk_2017_blind_C1 %>% 
  st_buffer(dist = 150)
trk_2017_main_C2_buff <- trk_2017 %>% 
  subset(TransLabel != "Baconian Swamp") %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)
trk_2017_blind_C2_buff <- trk_2017_blind_C2 %>% 
  st_buffer(dist = 150)

# bind all count 1 and count 2 together
trk_2017_C1_all_buff <- bind_rows(trk_2017_main_C1_buff, trk_2017_blind_C1_buff)
trk_2017_C2_all_buff <- bind_rows(trk_2017_main_C2_buff, trk_2017_blind_C2_buff)

# 2018
trk_2018_Bac_buff <- trk_2018 %>%
  subset(TransLabel == "Baconian Swamp") %>% 
  st_buffer(dist = 150)
trk_2018_main_C1_buff <- trk_2018 %>%
  subset(TransLabel != "Baconian Swamp") %>% 
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2018_main_C2_buff <- trk_2018 %>%
  subset(TransLabel != "Baconian Swamp") %>% 
  subset(Count == 2) %>% 
  st_buffer(dist = 150)

# 2019
trk_2019_buff <- trk_2019 %>% 
  st_buffer(dist = 150)

# 2020
trk_2020_Bac_C1_buff <- trk_2020 %>% 
  subset(TransLabel == "Baconian Swamp") %>%
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2020_Bac_C2_buff <- trk_2020 %>% 
  subset(TransLabel == "Baconian Swamp") %>%
  subset(Count == 2) %>% 
  st_buffer(dist = 150)
trk_2020_Cumbung2_C2_buff <- trk_2020 %>% 
  subset(TransLabel == "Cumbung 2") %>%
  subset(Count == 2) %>% 
  st_buffer(dist = 75) # one observer's data is missing here, so use half transect width than usual
trk_2020_main_C1_buff <- trk_2020 %>% 
  subset(TransLabel != "Baconian Swamp") %>%
  subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2020_main_C2_buff <- trk_2020 %>% 
  subset(TransLabel %in% c("Cumbung E1", "Lachlan L1", "Lower Lachlan 3", "Cumbung 3")) %>%
  subset(Count == 2) %>% 
  st_buffer(dist = 150)

# make new combo Baconian transects objects, as Lower Lachlan 3 is also located within Baconian Swamp
trk_2020_Bac_C1_all_buff <- trk_2020_main_C1_buff %>% 
  subset(TransLabel == "Lower Lachlan 3") %>% 
  bind_rows(., trk_2020_Bac_C1_buff)
trk_2020_Bac_C2_all_buff <- trk_2020_main_C2_buff %>% 
  subset(TransLabel == "Lower Lachlan 3") %>% 
  bind_rows(., trk_2020_Bac_C2_buff)

# also make new main Cumbung transects objects, without Lower Lachlan 3
trk_2020_main_only_C1_buff <- trk_2020_main_C1_buff %>% 
  subset(TransLabel != "Lower Lachlan 3") 
trk_2020_main_only_C2_buff <- trk_2020_main_C2_buff %>% 
  subset(TransLabel != "Lower Lachlan 3") %>% 
  bind_rows(., trk_2020_Cumbung2_C2_buff)


# 2021
trk_2021_Bac_buff <- trk_2021 %>% 
  subset(TransLabel == "Baconian Swamp") %>%
  st_buffer(dist = 150)
trk_2021_main_C1_buff <- trk_2021 %>% 
  subset(TransLabel != "Baconian Swamp") %>%
    subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2021_main_C2_buff <- trk_2021 %>% 
  subset(TransLabel != "Baconian Swamp") %>%
    subset(Count == 2) %>% 
  st_buffer(dist = 150)

# 2022 - all transects
trk_2022_buff <- trk_2022 %>% 
  st_buffer(dist = 100)

# 2023
trk_2023_Bac_buff <- trk_2023 %>% 
  subset(TransLabel == "Baconian") %>%
  st_buffer(dist = 150)
trk_2023_transect_buff <- trk_2023 %>% 
  subset(TransLabel != "Baconian") %>%
  subset(TransLabel != "Cumbung 1") %>%
  st_buffer(dist = 100)
trk_2023_Mean_C1_buff <- trk_2023 %>% 
  subset(TransLabel == "Cumbung 1") %>%
      subset(Count == 1) %>% 
  st_buffer(dist = 150)
trk_2023_Mean_C2_buff <- trk_2023 %>% 
  subset(TransLabel == "Cumbung 1") %>%
      subset(Count == 2) %>% 
  st_buffer(dist = 150)

# bind some buffered tracklogs together (all count 1 and all count 2) for calculations later
trk_2023_C1buff_all <- bind_rows(trk_2023_Mean_C1_buff,trk_2023_transect_buff)
plot(st_geometry(trk_2023_C1buff_all))

trk_2023_C2buff_all <- bind_rows(trk_2023_Mean_C2_buff,trk_2023_transect_buff)
plot(st_geometry(trk_2023_C2buff_all))


## 2024
trk_2024_Baconian_buff <- trk_2024_Baconian %>% 
  st_buffer(dist = 150)
plot(st_geometry(trk_2024_Baconian_buff))

trk_2024_C1_buff <- trk_2024_C1 %>% 
  st_buffer(dist = 150)

trk_2024_C2_buff <- trk_2024_C2 %>% 
  st_buffer(dist = 150)

trk_2024_transect_buff <- trk_2024_transect %>% 
  st_buffer(dist = 100)

# bind some buffered tracklogs together (all count 1 and all count 2) for calculations later
trk_2024_C1t_buff <- bind_rows(trk_2024_C1_buff,trk_2024_transect_buff)
plot(st_geometry(trk_2024_C1t_buff))

trk_2024_C2t_buff <- bind_rows(trk_2024_C2_buff,trk_2024_transect_buff)
plot(st_geometry(trk_2024_C2t_buff))

# 2025 - transects 1 and 2 are formal transects, and 3-6 are meandering
trk_2025_buff_transects <- trk_2025 %>%
  filter(name %in% c("Transect 1", "Transect 2")) %>% 
  st_buffer(dist = 100)
trk_2025_buff_meandering <- trk_2025 %>%
  filter(name %in% c("Transect 3", "Transect 4", "Transect 5", "Transect 6")) %>% 
  st_buffer(dist = 150)
# merge these datasets
trk_2025_buff <- bind_rows(trk_2025_buff_transects, trk_2025_buff_meandering)
plot(st_geometry(trk_2025_buff))
# these unfortunately overlap, need to deal with in the overlap section below.
```

Now, try plotting buffered tracklogs on their inundation maps.

```{r Plot_buffered_tracklogs}
# 2010

plot(inun_2010)
plot(trk_2010_buff, add = TRUE)
plot(trk_2010_AA_C1_buff, add = TRUE)
plot(trk_2010_AA_C2_buff, add = TRUE)

# 2011

plot(inun_2011)
plot(trk_2011_C1_all_buff, add = TRUE)
plot(trk_2011_C2_all_buff, add = TRUE)

# 2012

plot(inun_2012)
plot(trk_2012_C1_buff, add = TRUE)
plot(trk_2012_C2_buff, add = TRUE)

# 2013

plot(inun_2013)
plot(trk_2013_C1_buff, add = TRUE)
plot(trk_2013_C2_buff, add = TRUE)

# 2014

plot(inun_2014)
plot(trk_2014_C1_all_buff, add = TRUE)
plot(trk_2014_C2_all_buff, add = TRUE)

# 2015

plot(inun_2015)
plot(trk_2015_main_C1_buff, add = TRUE)
plot(trk_2015_main_C2_buff, add = TRUE)
plot(trk_2015_Bac_C1_buff, add = TRUE)
plot(trk_2015_Bac_C2_buff, add = TRUE)

# 2016

plot(inun_2016)
plot(trk_2016_transect_buff, add = TRUE)
plot(trk_20161109_C1_buff, add = TRUE)
plot(trk_20161109_C2_buff, add = TRUE)
plot(trk_20161107_Bac_C1_buff, add = TRUE)
plot(trk_20161107_Bac_C2_buff, add = TRUE)

# 2017

plot(inun_2017)
plot(trk_2017_C1_all_buff, add = TRUE)
plot(trk_2017_C2_all_buff, add = TRUE)
plot(trk_2017_Bac_buff, add = TRUE)

# 2018

plot(inun_2018)
plot(trk_2018_main_C1_buff, add = TRUE)
plot(trk_2018_main_C2_buff, add = TRUE)
plot(trk_2018_Bac_buff, add = TRUE)

# 2019

plot(inun_2019)
plot(trk_2019_buff, add = TRUE)

# 2020

plot(inun_2020)
plot(trk_2020_main_C1_buff, add = TRUE)
plot(trk_2020_main_C2_buff, add = TRUE)
plot(trk_2020_Bac_C1_buff, add = TRUE)
plot(trk_2020_Bac_C2_buff, add = TRUE)

# 2021

plot(inun_2021)
plot(trk_2021_main_C1_buff, add = TRUE)
plot(trk_2021_main_C2_buff, add = TRUE)
plot(trk_2021_Bac_buff, add = TRUE)

# 2022

plot(inun_2022)
plot(trk_2022_buff, add = TRUE)

# 2023

plot(inun_2023)
plot(trk_2023_Mean_C1_buff, add = TRUE)
plot(trk_2023_Mean_C2_buff, add = TRUE)
plot(trk_2023_Bac_buff, add = TRUE)
plot(trk_2023_transect_buff, add = TRUE)

# 2024

plot(inun_2024)
plot(trk_2024_C1_buff, add = TRUE)
plot(trk_2024_C2_buff, add = TRUE)
plot(trk_2024_Baconian_buff, add = TRUE)
plot(trk_2024_transect_buff, add = TRUE)

# 2025

plot(inun_2025)
plot(trk_2025_buff, add = TRUE)
```

## 3. Spatial manipulation and proportion calculation

Now, clip the inundation rasters by the buffered tracklogs to determine the inundated area surveyed by tracklogs. Also clip inundation rasters to cumbung ecological boundary and calculate the inundated area in this total boundary. Then calculate the proportion of inundated area surveyed out of the total inundated area.

Before this, rearrange our buffered transects into a sensible order. Right now, they're in a random order, which makes it more difficult to call specific transects.

Note, I haven't done this for every year so the other years are still in a non-arranged order and hence I've called haphazard numbers in the steps below. But I recommend you do this for all your buffered tracklogs!

```{r Reorder_buffered_transects}
trk_2010_buff <- trk_2010_buff %>% arrange(TransLabel)
```

Using different boundary files depending on if Baconian swamp and Marool lake are included as part of the cumbung ecological boundary for that year's survey:

-   2010: cumbung_eco

-   2011: cumbung_noMarool

-   2012: cumbung_noMarool

-   2013: cumbung_eco

-   2014: cumbung_noBac (even though no Baconian data, it was decided to include separately and Baconian as a zero count)

-   2015: cumbung_noBac_noM

-   2016: cumbung_noBac

-   2017: cumbung_noBac

-   2018: cumbung_noBac

-   2019: cumbung_eco

-   2020: cumbung_noBac

-   2021: cumbung_noBac_noM

-   2022: cumbung_eco

-   2023: cumbung_noBac

-   2024: cumbung_noBac_noM

-   2025: cumbung_noBac

For years with transect data, the output needed for input into TDW is an overall area and percentage inundated of the entire Great Cumbung Swamp, and an area and percentage inundated of each buffered tracklog. The names of each tracklog section are stored in the attribute tables of the shapefiles under the field "TransLabel".

Note, the function below this contains everything in a table, so feel free to skip this step.

```{r AreasProportions}
# Function to calculate area and proportion in a specific area
# input the year and name in parentheses
AreaProportionFn <- function(year,name,rasterf,boundary){
  y <- rasterf %>% 
  crop(extent(boundary)) %>% 
  mask(boundary) # clip the inundation raster to the cumbung boundary
  
  g <- as.data.frame(freq(y)) %>% subset(value == 1)
  g$area <- g$count # calculate no. cells inundated within the cumbung boundary
  
  print(paste0("In ", year, ", ", name, " area was ", round(st_area(boundary)/10000), " ha, and ", name, " inundated proportion was ", round(as.numeric(g$area*30*30/st_area(boundary) * 100),2), "%"))
  # calculate proportion inundated out of the whole area (noting that we need to multiply the number of cells in the inundated raster by 30x30 as that's the size of the cells)
  }


# 2024
AreaProportionFn("2024", "Overall wetland", inun_2024,cumbung_noBac_noM)
AreaProportionFn("2024", "Baconian Swamp", inun_2024,Baconian)
AreaProportionFn("2024", "Baconian Swamp transect", inun_2024, trk_2024_Baconian_buff)
AreaProportionFn("2024", "Cumbung T1", inun_2024, trk_2024_transect_buff[7,])
AreaProportionFn("2024", "Cumbung T2", inun_2024, trk_2024_transect_buff[8,])
AreaProportionFn("2024", "Cumbung T3", inun_2024, trk_2024_transect_buff[9,])
AreaProportionFn("2024", "Cumbung T4", inun_2024, trk_2024_transect_buff[10,])
AreaProportionFn("2024", "Cumbung T5", inun_2024, trk_2024_transect_buff[1,])
AreaProportionFn("2024", "Cumbung T6", inun_2024, trk_2024_transect_buff[2,])
AreaProportionFn("2024", "Cumbung T7", inun_2024, trk_2024_transect_buff[3,])
AreaProportionFn("2024", "Cumbung T8", inun_2024, trk_2024_transect_buff[4,])
AreaProportionFn("2024", "Cumbung T9", inun_2024, trk_2024_transect_buff[5,])
AreaProportionFn("2024", "Cumbung T10", inun_2024, trk_2024_transect_buff[6,])
AreaProportionFn("2024", "Count 1 Cumbung 1", inun_2024, trk_2024_C1_buff[1,])
AreaProportionFn("2024", "Count 1 Cumbung North 3", inun_2024, trk_2024_C1_buff[2,])
AreaProportionFn("2024", "Count 1 Cumbung Lake", inun_2024, trk_2024_C1_buff[3,])
AreaProportionFn("2024", "Count 2 Cumbung 1", inun_2024, trk_2024_C2_buff[1,])
AreaProportionFn("2024", "Count 2 Cumbung Lake", inun_2024, trk_2024_C2_buff[2,])
```

Table output.

```{r Output_table}
# buffboundary is the buffered tracklog. For overall wetland calculations, put the overall wetland boundary for both boundary and buffboundary
AreaPropString <- function(name, rasterf, boundary, buffboundary){
  y <- rasterf %>% 
  crop(extent(boundary)) %>% 
  mask(boundary) %>%  # clip the inundation raster to the cumbung boundary
  mask(buffboundary) # also clip to the boundary of the buffered tracklog, where applicable 
  
  g <- as.data.frame(freq(y)) %>% subset(value == 1) 
  
  # Check if g is empty and assign 0 if it is
  if (nrow(g) == 0) {
    g <- data.frame(value = 1, count = 0)
  }
  
  g$area <- g$count # calculate no. cells inundated within the cumbung boundary
  
  a <- round(st_area(buffboundary)/10000)
  
  p <- round(as.numeric(g$area*30*30/st_area(buffboundary) * 100),2)
 
  # assign to string
  print(c(name, a, p))
}

# modified version of the above function for rasters with 10x10m cells (the above is for inundation rasters with 30x30m cells)
AreaPropString10 <- function(name, rasterf, boundary, buffboundary){
  y <- rasterf %>% 
  crop(extent(boundary)) %>% 
  mask(boundary) %>%  # clip the inundation raster to the cumbung boundary
  mask(buffboundary) # also clip to the boundary of the buffered tracklog, where applicable 
  
  g <- as.data.frame(freq(y)) %>% subset(value == 1) 
  
  # Check if g is empty and assign 0 if it is
  if (nrow(g) == 0) {
    g <- data.frame(value = 1, count = 0)
  }
  
  g$area <- g$count # calculate no. cells inundated within the cumbung boundary
  
  a <- round(st_area(buffboundary)/10000)
  
  p <- round(as.numeric(g$area*10*10/st_area(buffboundary) * 100),2)
 
  # assign to string
  print(c(name, a, p))
}



# 2010
Stats2010 <- matrix(c(AreaPropString("Overall wetland", inun_2010, cumbung_eco, cumbung_eco),
                      AreaPropString("Cumbung Transect 1", inun_2010, cumbung_eco, trk_2010_buff[1,]),
                      AreaPropString("Cumbung Transect 2", inun_2010, cumbung_eco, trk_2010_buff[2,]),
                      AreaPropString("Cumbung Transect 3", inun_2010, cumbung_eco, trk_2010_buff[3,]),
                      AreaPropString("Cumbung Transect 4", inun_2010, cumbung_eco, trk_2010_buff[4,]),
                      AreaPropString("Cumbung Transect 5", inun_2010, cumbung_eco, trk_2010_buff[5,]),
                      AreaPropString("Cumbung Transect 6", inun_2010, cumbung_eco, trk_2010_buff[6,]),
                      AreaPropString("Cumbung Transect 7", inun_2010, cumbung_eco, trk_2010_buff[7,]),
                      AreaPropString("Cumbung AA Count 1", inun_2010, cumbung_eco, trk_2010_AA_C1_buff),
                      AreaPropString("Cumbung AA Count 2", inun_2010, cumbung_eco, trk_2010_AA_C2_buff)),
               ncol = 3, byrow = TRUE)
colnames(Stats2010) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')



# 2011
Stats2011 <- matrix(c(AreaPropString("Overall wetland", inun_2011, cumbung_noMarool, cumbung_noMarool),
                      AreaPropString("Cumbung Swamp Count 1", inun_2011, cumbung_noMarool, trk_2011_C1_all_buff[5,]),
                      AreaPropString("Lachlan River Count 1", inun_2011, cumbung_noMarool, trk_2011_C1_all_buff[4,]),
                      AreaPropString("Cumbung A6 Count 1", inun_2011, cumbung_noMarool, trk_2011_C1_all_buff[3,]),
                      AreaPropString("Cumbung A7 Count 1", inun_2011, cumbung_noMarool, trk_2011_C1_all_buff[2,]),
                      AreaPropString("Cumbung A8 Count 1", inun_2011, cumbung_noMarool, trk_2011_C1_all_buff[1,]),
                      AreaPropString("Blindbungi Count 1", inun_2011, cumbung_noMarool, trk_2011_C1_all_buff[6,]),
                      AreaPropString("Cumbung Swamp Count 2", inun_2011, cumbung_noMarool, trk_2011_C2_all_buff[5,]),
                      AreaPropString("Lachlan River Count 2", inun_2011, cumbung_noMarool, trk_2011_C2_all_buff[4,]),
                      AreaPropString("Cumbung A6 Count 2", inun_2011, cumbung_noMarool, trk_2011_C2_all_buff[3,]),
                      AreaPropString("Cumbung A7 Count 2", inun_2011, cumbung_noMarool, trk_2011_C2_all_buff[2,]),
                      AreaPropString("Cumbung A8 Count 2", inun_2011, cumbung_noMarool, trk_2011_C2_all_buff[1,]),
                      AreaPropString("Blindbungi Count 2", inun_2011, cumbung_noMarool, trk_2011_C2_all_buff[6,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2011) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2012
Stats2012 <- matrix(c(AreaPropString("Overall wetland", inun_2012, cumbung_noMarool, cumbung_noMarool),
                      AreaPropString("Cumbung S1 Count 1", inun_2012, cumbung_noMarool, trk_2012_C1_buff[1,]),
                      AreaPropString("Cumbung 1 Count 1", inun_2012, cumbung_noMarool, trk_2012_C1_buff[2,]),
                      AreaPropString("Cumbung S2 Count 1", inun_2012, cumbung_noMarool, trk_2012_C1_buff[3,]),
                      AreaPropString("Cumbung 4a Count 1", inun_2012, cumbung_noMarool, trk_2012_C1_buff[4,]),
                      AreaPropString("Cumbung S1 Count 2", inun_2012, cumbung_noMarool, trk_2012_C2_buff[1,]),
                      AreaPropString("Cumbung 1 Count 2", inun_2012, cumbung_noMarool, trk_2012_C2_buff[2,]),
                      AreaPropString("Cumbung S2 Count 2", inun_2012, cumbung_noMarool, trk_2012_C2_buff[3,]),
                      AreaPropString("Cumbung 4a Count 2", inun_2012, cumbung_noMarool, trk_2012_C2_buff[4,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2012) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2013
Stats2013 <- matrix(c(AreaPropString("Overall wetland", inun_2013, cumbung_eco, cumbung_eco),
                      AreaPropString("Cumbung Clear Lake Count 1", inun_2013, cumbung_eco, trk_2013_C1_buff[1,]),
                      AreaPropString("Cumbung North Count 1", inun_2013, cumbung_eco, trk_2013_C1_buff[2,]),
                      AreaPropString("Cumbung AA Count 1", inun_2013, cumbung_eco, trk_2013_C1_buff[3,]),
                      AreaPropString("Cumbung North Count 2", inun_2013, cumbung_eco, trk_2013_C2_buff[2,]),
                      AreaPropString("Cumbung AA Count 2", inun_2013, cumbung_eco, trk_2013_C2_buff[1,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2013) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2014
Stats2014 <- matrix(c(AreaPropString("Overall wetland", inun_2014, cumbung_noBac, cumbung_eco),
                      AreaPropString("Cumbung Swamp Count 1", inun_2014, cumbung_noBac, trk_2014_C1_all_buff[1,]),
                      AreaPropString("Cumbung Swamp Count 2", inun_2014, cumbung_noBac, trk_2014_C2_all_buff[1,]),
                      AreaPropString("Blindbungi Count 1", inun_2014, cumbung_noBac, trk_2014_C1_all_buff[2,]),
                      AreaPropString("Blindbungi Count 2", inun_2014, cumbung_noBac, trk_2014_C2_all_buff[2,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2014) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2015
Stats2015 <- matrix(c(AreaPropString("Overall wetland", inun_2015, cumbung_noBac_noM, cumbung_noBac_noM),
                      AreaPropString("Baconian Swamp overall", inun_2015, Baconian, Baconian),
                      AreaPropString("Baconian Swamp Count 1", inun_2015, Baconian, trk_2015_Bac_C1_buff),
                      AreaPropString("Baconian Swamp Count 2", inun_2015, Baconian, trk_2015_Bac_C2_buff),
                      AreaPropString("Cumbung S Count 1", inun_2015, cumbung_noBac_noM, trk_2015_main_C1_buff[1,]),
                      AreaPropString("Cumbung Swamp Count 1", inun_2015, cumbung_noBac_noM, trk_2015_main_C1_buff[2,]),
                      AreaPropString("Cumbung B Count 1", inun_2015, cumbung_noBac_noM, trk_2015_main_C1_buff[3,]),
                      AreaPropString("Cumbung AA Count 1", inun_2015, cumbung_noBac_noM, trk_2015_main_C1_buff[4,]),
                      AreaPropString("Cumbung S Count 2", inun_2015, cumbung_noBac_noM, trk_2015_main_C2_buff[1,]),
                      AreaPropString("Cumbung Swamp Count 2", inun_2015, cumbung_noBac_noM, trk_2015_main_C2_buff[3,]),
                      AreaPropString("Cumbung B Count 2", inun_2015, cumbung_noBac_noM, trk_2015_main_C2_buff[4,]),
                      AreaPropString("Cumbung AA Count 2", inun_2015, cumbung_noBac_noM, trk_2015_main_C2_buff[2,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2015) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')

# 2016
Stats2016 <- matrix(c(AreaPropString("Overall wetland", inun_2016, cumbung_noBac, cumbung_noBac),
                      AreaPropString("Baconian Swamp overall", inun_2016, Baconian, Baconian),
                      AreaPropString("Baconian Swamp Count 1", inun_2016, Baconian, trk_20161107_Bac_C1_buff),
                      AreaPropString("Baconian Swamp Count 2", inun_2016, Baconian, trk_20161107_Bac_C2_buff),
                      AreaPropString("Cumbung Transect 1", inun_2016, cumbung_noBac, trk_2016_transect_buff[6,]),
                      AreaPropString("Cumbung Transect 2", inun_2016, cumbung_noBac, trk_2016_transect_buff[5,]),
                      AreaPropString("Cumbung Transect 3", inun_2016, cumbung_noBac, trk_2016_transect_buff[4,]),
                      AreaPropString("Cumbung Transect 4", inun_2016, cumbung_noBac, trk_2016_transect_buff[3,]),
                      AreaPropString("Cumbung Transect 5", inun_2016, cumbung_noBac, trk_2016_transect_buff[2,]),
                      AreaPropString("Cumbung Transect 6", inun_2016, cumbung_noBac, trk_2016_transect_buff[7,]),
                      AreaPropString("Cumbung Transect 7", inun_2016, cumbung_noBac, trk_2016_transect_buff[8,]),
                      AreaPropString("Cumbung Transect 8", inun_2016, cumbung_noBac, trk_2016_transect_buff[1,]),
                      AreaPropString("Cumbung Transect 9", inun_2016, cumbung_noBac, trk_2016_transect_buff[9,]),
                      AreaPropString("Cumbung Transect 10", inun_2016, cumbung_noBac, trk_2016_transect_buff[10,]),
                      AreaPropString("Cumbung 3 Count 1", inun_2016, cumbung_noBac, trk_20161109_C1_buff[1,]),
                      AreaPropString("Cumbung 3A Count 1", inun_2016, cumbung_noBac, trk_20161109_C1_buff[2,]),
                      AreaPropString("Cumbung 3 Count 2", inun_2016, cumbung_noBac, trk_20161109_C2_buff[2,]),
                      AreaPropString("Cumbung 3A Count 2", inun_2016, cumbung_noBac, trk_20161109_C2_buff[1,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2016) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')



# 2017
Stats2017 <- matrix(c(AreaPropString("Overall wetland", inun_2017, cumbung_noBac, cumbung_noBac),
                      AreaPropString("Baconian Swamp overall", inun_2017, Baconian, Baconian),
                      AreaPropString("Baconian Swamp", inun_2017, Baconian, trk_2017_Bac_buff),
                      AreaPropString("Cumbung 1 Count 1", inun_2017, cumbung_noBac, trk_2017_C1_all_buff[1,]),
                      AreaPropString("Lignum Lake (not) Count 1", inun_2017, cumbung_noBac, trk_2017_C1_all_buff[2,]),
                      AreaPropString("Cumbung 3 Count 1", inun_2017, cumbung_noBac, trk_2017_C1_all_buff[3,]),
                      AreaPropString("Blindbungi Count 1", inun_2017, cumbung_noBac, trk_2017_C1_all_buff[4,]),
                      AreaPropString("Cumbung 1 Count 2", inun_2017, cumbung_noBac, trk_2017_C2_all_buff[1,]),
                      AreaPropString("Cumbung 3 Count 2", inun_2017, cumbung_noBac, trk_2017_C2_all_buff[2,]),
                      AreaPropString("Blindbungi Count 2", inun_2017, cumbung_noBac, trk_2017_C2_all_buff[3,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2017) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')

# 2018
Stats2018 <- matrix(c(AreaPropString("Overall wetland", inun_2018, cumbung_noBac, cumbung_noBac),
                      AreaPropString("Baconian Swamp overall", inun_2018, Baconian, Baconian),
                      AreaPropString("Baconian Swamp", inun_2018, Baconian, trk_2018_Bac_buff),
                      AreaPropString("Great Cumbung Swamp 3 (1) Count 1", inun_2018, cumbung_noBac, trk_2018_main_C1_buff[1,]),
                      AreaPropString("Great Cumbung Swamp 3 (2) Count 1", inun_2018, cumbung_noBac, trk_2018_main_C1_buff[2,]),
                      AreaPropString("Cumbung 2 Count 1", inun_2018, cumbung_noBac, trk_2018_main_C1_buff[3,]),
                      AreaPropString("Great Cumbung Swamp 3 (1) Count 2", inun_2018, cumbung_noBac, trk_2018_main_C2_buff[1,]),
                      AreaPropString("Great Cumbung Swamp 3 (2) Count 2", inun_2018, cumbung_noBac, trk_2018_main_C2_buff[3,]),
                      AreaPropString("Cumbung 2 Count 2", inun_2018, cumbung_noBac, trk_2018_main_C2_buff[2,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2018) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')

# 2019
# Note, 2019 has a raster with 10x10m cells rather than 30x30m cells. Use a modified version of the function for 10x10m cells
Stats2019 <- matrix(c(AreaPropString10("Overall wetland", inun_2019, cumbung_eco, cumbung_eco),
                      AreaPropString10("Cumbung 3 Count 1", inun_2019, cumbung_eco, trk_2019_buff[2,]),
                      AreaPropString10("Cumbung 4 Count 1", inun_2019, cumbung_eco, trk_2019_buff[1,]),
                      AreaPropString10("Great Cumbung Swamp Count 1", inun_2019, cumbung_eco, trk_2019_buff[3,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2019) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2020
# Note, 2020 has a raster with 10x10m cells rather than 30x30m cells. Use a modified version of the function for 10x10m cells
# Lower Lachlan 3 is in Baconian so therefore its proportion relates to Baconian
Stats2020 <- matrix(c(AreaPropString10("Overall wetland", inun_2020, cumbung_noBac, cumbung_noBac),
                      AreaPropString10("Baconian Swamp overall", inun_2020, Baconian, Baconian),
                      AreaPropString10("Baconian Swamp Count 1", inun_2020, Baconian, trk_2020_Bac_C1_all_buff[2,]),
                      AreaPropString10("Baconian Swamp Count 2", inun_2020, Baconian, trk_2020_Bac_C2_all_buff[2,]),
                      AreaPropString10("Lower Lachlan 3 Count 1", inun_2020, Baconian, trk_2020_Bac_C1_all_buff[1,]),
                       AreaPropString10("Lower Lachlan 3 Count 2", inun_2020, Baconian, trk_2020_Bac_C2_all_buff[1,]),
                      AreaPropString10("Cumbung E1 Count 1", inun_2020, cumbung_noBac, trk_2020_main_only_C1_buff[1,]),
                      AreaPropString10("Cumbung 2 Count 1", inun_2020, cumbung_noBac, trk_2020_main_only_C1_buff[2,]),
                      AreaPropString10("Lachlan L1 Count 1", inun_2020, cumbung_noBac, trk_2020_main_only_C1_buff[3,]),
                      AreaPropString10("Cumbung 3 Count 1", inun_2020, cumbung_noBac, trk_2020_main_only_C1_buff[4,]),
                      AreaPropString10("Cumbung E1 Count 2", inun_2020, cumbung_noBac, trk_2020_main_only_C2_buff[1,]),
                      AreaPropString10("Cumbung 2 Count 2", inun_2020, cumbung_noBac, trk_2020_main_only_C2_buff[4,]),
                      AreaPropString10("Lachlan L1 Count 2", inun_2020, cumbung_noBac, trk_2020_main_only_C2_buff[2,]),
                      AreaPropString10("Cumbung 3 Count 2", inun_2020, cumbung_noBac, trk_2020_main_only_C2_buff[3,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2020) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2021
# Note, 2021 has a raster with 10x10m cells rather than 30x30m cells. Use a modified version of the function for 10x10m cells
Stats2021 <- matrix(c(AreaPropString10("Overall wetland", inun_2021, cumbung_noBac_noM, cumbung_noBac_noM),
                      AreaPropString10("Baconian Swamp overall", inun_2021, Baconian, Baconian),
                      AreaPropString10("Baconian Swamp", inun_2021, Baconian, trk_2021_Bac_buff),
                      AreaPropString10("Cumbung AA Count 1", inun_2021, cumbung_noBac_noM, trk_2021_main_C1_buff[1,]),
                      AreaPropString10("Boyong Lake (not) Count 1", inun_2021, cumbung_noBac_noM, trk_2021_main_C1_buff[2,]),
                      AreaPropString10("Great Cumbung Swamp Count 1", inun_2021, cumbung_noBac_noM, trk_2021_main_C1_buff[3,]),
                      AreaPropString10("Cumbung AA Count 2", inun_2021, cumbung_noBac_noM, trk_2021_main_C2_buff[3,]),
                      AreaPropString10("Boyong Lake (not) Count 2", inun_2021, cumbung_noBac_noM, trk_2021_main_C2_buff[2,]),
                      AreaPropString10("Great Cumbung Swamp Count 2", inun_2021, cumbung_noBac_noM, trk_2021_main_C2_buff[1,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2021) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2022
# Note, some of these transects are certainly 100% inundated. The calculations sometimes have them slightly above 100% - this is because they're based on 30 x 30 m cell sizes so the area stats are not super precise.
Stats2022 <- matrix(c(AreaPropString("Overall wetland", inun_2022, cumbung_eco, cumbung_eco),
                      AreaPropString("Cumbung 1", inun_2022, cumbung_eco, trk_2022_buff[4,]),
                      AreaPropString("Cumbung 2", inun_2022, cumbung_eco, trk_2022_buff[6,]),
                      AreaPropString("Cumbung 3", inun_2022, cumbung_eco, trk_2022_buff[9,]),
                      AreaPropString("Cumbung 4", inun_2022, cumbung_eco, trk_2022_buff[7,]),
                      AreaPropString("Cumbung 5", inun_2022, cumbung_eco, trk_2022_buff[1,]),
                      AreaPropString("Cumbung 6", inun_2022, cumbung_eco, trk_2022_buff[10,]),
                      AreaPropString("Cumbung 7", inun_2022, cumbung_eco, trk_2022_buff[5,]),
                      AreaPropString("Cumbung 8", inun_2022, cumbung_eco, trk_2022_buff[2,]),
                      AreaPropString("Cumbung 9", inun_2022, cumbung_eco, trk_2022_buff[8,]),
                      AreaPropString("Cumbung 10", inun_2022, cumbung_eco, trk_2022_buff[3,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2022) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')

# 2023
Stats2023 <- matrix(c(AreaPropString("Overall wetland", inun_2023, cumbung_noBac, cumbung_noBac),
                      AreaPropString("Baconian Swamp overall", inun_2023, Baconian, Baconian),
                      AreaPropString("Baconian", inun_2023, Baconian, trk_2023_Bac_buff),
                      AreaPropString("Cumbung 1 Count 1", inun_2023, cumbung_noBac, trk_2023_Mean_C1_buff),
                      AreaPropString("Cumbung 1 Count 2", inun_2023, cumbung_noBac, trk_2023_Mean_C2_buff),
                      AreaPropString("Great Cumbung Swamp Transect 1", inun_2023, cumbung_noBac, trk_2023_transect_buff[1,]),
                      AreaPropString("Great Cumbung Swamp Transect 2", inun_2023, cumbung_noBac, trk_2023_transect_buff[2,]),
                      AreaPropString("Great Cumbung Swamp Transect 3", inun_2023, cumbung_noBac, trk_2023_transect_buff[3,]),
                      AreaPropString("Great Cumbung Swamp Transect 4", inun_2023, cumbung_noBac, trk_2023_transect_buff[4,]),
                      AreaPropString("Great Cumbung Swamp Transect 5", inun_2023, cumbung_noBac, trk_2023_transect_buff[5,]),
                      AreaPropString("Great Cumbung Swamp Transect 6", inun_2023, cumbung_noBac, trk_2023_transect_buff[6,]),
                      AreaPropString("Great Cumbung Swamp Transect 7", inun_2023, cumbung_noBac, trk_2023_transect_buff[7,]),
                      AreaPropString("Great Cumbung Swamp Transect 8", inun_2023, cumbung_noBac, trk_2023_transect_buff[9,]),
                      AreaPropString("Great Cumbung Swamp Transect 9", inun_2023, cumbung_noBac, trk_2023_transect_buff[10,]),
                      AreaPropString("Great Cumbung Swamp Transect 10", inun_2023, cumbung_noBac, trk_2023_transect_buff[8,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2023) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2024

Stats2024 <- matrix(c(AreaPropString("Overall wetland", inun_2024,cumbung_noBac_noM,cumbung_noBac_noM),
  AreaPropString("Baconian Swamp", inun_2024,Baconian,Baconian),
  AreaPropString("Baconian Swamp transect", inun_2024,Baconian, trk_2024_Baconian_buff),
                 AreaPropString("Cumbung T1", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[7,]),
               AreaPropString("Cumbung T2", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[8,]),
               AreaPropString("Cumbung T3", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[9,]),
               AreaPropString("Cumbung T4", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[10,]),
               AreaPropString("Cumbung T5", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[1,]),
               AreaPropString("Cumbung T6", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[2,]),
               AreaPropString("Cumbung T7", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[3,]),
               AreaPropString("Cumbung T8", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[4,]),
               AreaPropString("Cumbung T9", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[5,]),
               AreaPropString("Cumbung T10", inun_2024,cumbung_noBac_noM, trk_2024_transect_buff[6,]),
               AreaPropString("Count 1 Cumbung 1", inun_2024,cumbung_noBac_noM, trk_2024_C1_buff[1,]),
               AreaPropString("Count 1 Cumbung North 3", inun_2024,cumbung_noBac_noM, trk_2024_C1_buff[2,]),
               AreaPropString("Count 1 Cumbung Lake", inun_2024,cumbung_noBac_noM, trk_2024_C1_buff[3,]),
               AreaPropString("Count 2 Cumbung 1", inun_2024,cumbung_noBac_noM, trk_2024_C2_buff[1,]),
               AreaPropString("Count 2 Cumbung Lake", inun_2024,cumbung_noBac_noM, trk_2024_C2_buff[2,])),
               ncol = 3, byrow = TRUE)

colnames(Stats2024) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')


# 2025
Stats2025 <- matrix(c(AreaPropString("Overall wetland", inun_2025, cumbung_noBac, cumbung_noBac),
                      AreaPropString("Transect 1", inun_2025, cumbung_noBac, trk_2025_buff[2,]),
                      AreaPropString("Transect 2", inun_2025, cumbung_noBac, trk_2025_buff[1,]),
                      AreaPropString("Transect 3", inun_2025, cumbung_noBac, trk_2025_buff[3,]),
                      AreaPropString("Transect 4", inun_2025, cumbung_noBac, trk_2025_buff[4,]),
                      AreaPropString("Transect 5", inun_2025, cumbung_noBac, trk_2025_buff[5,]),
                      AreaPropString("Transect 6", inun_2025, cumbung_noBac, trk_2025_buff[6,])),
               ncol = 3, byrow = TRUE)
colnames(Stats2025) <- c('Transect name', 'Area (ha)', 'Proportion inundated (%)')
```

The following chunk calculates the overall percentage of inundated area in the total wetland that was surveyed in all tracklogs together. This is the input in TDW for years without transect data and should be used as a quality check in years with transect data.

```{r Inundation_areas}
# input the year, location and count number (countno -- 1 or 2) in parentheses
inun_prop_surveyed <- function(year,location,countno,rasterf,tracklog,boundary){
  x <- rasterf %>% 
  crop(extent(tracklog)) %>% 
  mask(tracklog) # clip the inundation raster to the tracklog
  
  f <- as.data.frame(freq(x)) %>% subset(value == 1)
  
  # Check if f is empty and assign 0 if it is
  if (nrow(f) == 0) {
    f <- data.frame(value = 1, count = 0)
  }
  
  f$area <- f$count # calculate no. cells inundated within the tracklog area
  
  y <- rasterf %>% 
  crop(extent(boundary)) %>% 
  mask(boundary) # clip the inundation raster to the cumbung boundary
  
  g <- as.data.frame(freq(y)) %>% subset(value == 1)
  g$area <- g$count # calculate no. cells inundated within the cumbung boundary
  
  print(paste0("In ", year, ", at ", location, " in count ", countno, ", the proportion of total inundated area surveyed was ", round(f$area /g$area * 100, 2), "%")) # now calculate what percentage of the total (cumbung) inundated area was surveyed within the tracklog area
}

# 2010
Prop2010C1 <- inun_prop_surveyed("2010", "Cumbung Swamp", "1", inun_2010, trk_2010_C1_all_buff, cumbung_eco)
Prop2010C2 <- inun_prop_surveyed("2010", "Cumbung Swamp", "2", inun_2010, trk_2010_C2_all_buff, cumbung_eco)

# 2011
Prop2011C1 <- inun_prop_surveyed("2011", "Cumbung Swamp", "1", inun_2011, trk_2011_C1_all_buff, cumbung_noMarool)
Prop2011C2 <- inun_prop_surveyed("2011", "Cumbung Swamp", "2", inun_2011, trk_2011_C2_all_buff, cumbung_noMarool)

# 2012
Prop2012C1 <- inun_prop_surveyed("2012", "Cumbung Swamp", "1", inun_2012, trk_2012_C1_buff, cumbung_noMarool)
Prop2012C2 <- inun_prop_surveyed("2012", "Cumbung Swamp", "2", inun_2012, trk_2012_C2_buff, cumbung_noMarool)

# 2013
Prop2013C1 <- inun_prop_surveyed("2013", "Cumbung Swamp", "1", inun_2013, trk_2013_C1_buff, cumbung_eco)
Prop2013C2 <- inun_prop_surveyed("2013", "Cumbung Swamp", "2", inun_2013, trk_2013_C2_buff, cumbung_eco)

# 2014
Prop2014C1 <- inun_prop_surveyed("2014", "Cumbung Swamp", "1", inun_2014, trk_2014_C1_all_buff, cumbung_noBac)
Prop2014C2 <- inun_prop_surveyed("2014", "Cumbung Swamp", "2", inun_2014, trk_2014_C2_all_buff, cumbung_noBac)


# 2015
Prop2015BacC1 <- inun_prop_surveyed("2015", "Baconian Swamp", "1", inun_2015, trk_2015_Bac_C1_buff, Baconian)
Prop2015BacC2 <- inun_prop_surveyed("2015", "Baconian Swamp", "2", inun_2015, trk_2015_Bac_C2_buff, Baconian)
Prop2015C1 <- inun_prop_surveyed("2015", "Cumbung Swamp", "1", inun_2015, trk_2015_main_C1_buff, cumbung_noBac_noM)
Prop2015C2 <- inun_prop_surveyed("2015", "Cumbung Swamp", "2", inun_2015, trk_2015_main_C2_buff, cumbung_noBac_noM)

# 2016
Prop20161109C1 <- inun_prop_surveyed("2016, on 9 November", "Cumbung Swamp", "1", inun_2016, trk_20161109_C1_all, cumbung_noBac)
Prop20161109C2 <- inun_prop_surveyed("2016, on 9 November", "Cumbung Swamp", "2", inun_2016, trk_20161109_C2_all, cumbung_noBac)
Prop20161107BacC1 <- inun_prop_surveyed("2016, on 7 November", "Baconian Swamp", "1", inun_2016, trk_20161107_Bac_C1_buff, Baconian)
Prop20161107BacC2 <- inun_prop_surveyed("2016, on 7 November", "Baconian Swamp", "2", inun_2016, trk_20161107_Bac_C2_buff, Baconian)

# 2017
Prop2017Bac <- inun_prop_surveyed("2017", "Baconian Swamp", "1", inun_2017, trk_2017_Bac_buff, Baconian)
Prop2017C1 <- inun_prop_surveyed("2017", "Cumbung Swamp", "1", inun_2017, trk_2017_C1_all_buff, cumbung_noBac)
Prop2017C2 <- inun_prop_surveyed("2017", "Cumbung Swamp", "2", inun_2017, trk_2017_C2_all_buff, cumbung_noBac)

# 2018
Prop2018Bac <- inun_prop_surveyed("2018", "Baconian Swamp", "1", inun_2018, trk_2018_Bac_buff, Baconian)
Prop2018C1 <- inun_prop_surveyed("2018", "Cumbung Swamp", "1", inun_2018, trk_2018_main_C1_buff, cumbung_noBac)
Prop2018C2 <- inun_prop_surveyed("2018", "Cumbung Swamp", "2", inun_2018, trk_2018_main_C2_buff, cumbung_noBac)

# 2019
Prop2019 <- inun_prop_surveyed("2019", "Cumbung Swamp", "1", inun_2019, trk_2019_buff, cumbung_eco)

# 2020
Prop2020BacC1 <- inun_prop_surveyed("2020", "Baconian Swamp", "1", inun_2020, trk_2020_Bac_C1_all_buff, Baconian)
Prop2020BacC2 <- inun_prop_surveyed("2020", "Baconian Swamp", "2", inun_2020, trk_2020_Bac_C2_all_buff, Baconian)
Prop2020C1 <- inun_prop_surveyed("2020", "Cumbung Swamp", "1", inun_2020, trk_2020_main_only_C1_buff, cumbung_noBac)
Prop2020C2 <- inun_prop_surveyed("2020", "Cumbung Swamp", "2", inun_2020, trk_2020_main_only_C2_buff, cumbung_noBac)

# 2021
Prop2021Bac <- inun_prop_surveyed("2021", "Baconian Swamp", "1", inun_2021, trk_2021_Bac_buff, Baconian)
Prop2021C1 <- inun_prop_surveyed("2021", "Cumbung Swamp", "1", inun_2021, trk_2021_main_C1_buff, cumbung_noBac_noM)
Prop2021C2 <- inun_prop_surveyed("2021", "Cumbung Swamp", "2", inun_2021, trk_2021_main_C2_buff, cumbung_noBac_noM)

# 2022
Prop2022 <- inun_prop_surveyed("2022", "Cumbung Swamp", "1", inun_2022, trk_2022_buff, cumbung_eco)

# 2023
Prop2023Bac <- inun_prop_surveyed("2023", "Baconian Swamp", "1", inun_2023, trk_2023_Bac_buff, Baconian)
Prop2023C1 <- inun_prop_surveyed("2023", "Cumbung Swamp", "1", inun_2023, trk_2023_C1buff_all, cumbung_noBac)
Prop2023C1 <- inun_prop_surveyed("2023", "Cumbung Swamp", "2", inun_2023, trk_2023_C2buff_all, cumbung_noBac)

# 2024
Prop2024Bac <- inun_prop_surveyed("2024", "Baconian Swamp", "1", inun_2024,  trk_2024_Baconian_buff, Baconian)
Prop_2024C1 <- inun_prop_surveyed("2024", "Cumbung Swamp", "1", inun_2024, trk_2024_C1t_buff, cumbung_noBac_noM)
Prop_2024C2 <- inun_prop_surveyed("2024", "Cumbung Swamp", "2", inun_2024, trk_2024_C2t_buff, cumbung_noBac_noM)

# 2025
Prop2025 <- inun_prop_surveyed("2025", "Cumbung Swamp", "1", inun_2025, trk_2025_buff, cumbung_noBac)
```

## 4. Overlap areas

**Calculate overlapped tracklog buffer areas**

In some years, buffered tracklogs overlap with one another (2010, 2017, 2018, 2019). To avoid double counting these areas in The Duck Web, we need to calculate the area that is overlapped in each case and subtract half from each buffered tracklog area.

-   In 2010, track 4 & 5 overlap one another (lines 7 & 4 in trk_2010_buff)
-   In 2016:
    -   track 6 and 7 overlap (lines 9 & 10 in trk_20161109_C1_all)
    -   cumbung 3A overlaps with track 9 and track 8 (lines 2, 3 and 11)
-   In 2017, Cumbung 3 and Lignum Lake Not overlap (lines 3 and 2 in trk_2017_C1_all)
-   In 2018, GCS and GCS2 overlap (lines 1 and 2 in trk_2018_main_C1_buff)
-   In 2019, Cumbung 3 and GCS overlap (lines 2 and 3 in trk_2019_buff)
-   In 2020, Cumbung E1 and Cumbung 3 overlap (lines 1 and 5 in trk_2020_main_C1_buff)
-   In 2025, transects 3 & 5 overlap, and transects 4 & 6 overlap.

Calculate the area of each overlapping buffered tracklog individually, and then their total area, to determine the area of the overlap.

```{r Overlapping_tracklogs_2010}
#2010

# first, visualise transects
plot(inun_2010)
plot(trk_2010_buff, add = TRUE)
plot(trk_2010_AA_C1_buff, add = TRUE)
#plot(trk_2010_AA_C2_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2010_buff_T4_5_inun <- inun_2010 %>% 
  crop(extent(trk_2010_buff[c(4,7),])) %>% 
  mask(trk_2010_buff[c(4,7),])

area_2010_T4_5 <- as.data.frame(freq(trk_2010_buff_T4_5_inun)) 
area_2010_T4_5_inun <- area_2010_T4_5 %>% subset(value == 1)
area_2010_T4_5_noninun <- area_2010_T4_5 %>% subset(value == 0)

# overlapped area = track 4 area + track 5 area - track4&5 area
plot(trk_2010_buff_T4_5_inun)

overlaparea2010 <- as.numeric(st_area(trk_2010_buff[7,])/10000) + 
  as.numeric(st_area(trk_2010_buff[4,])/10000) - 
  (area_2010_T4_5_inun$count + area_2010_T4_5_noninun$count)*30*30/10000 # account for cell size

# overlapped inundation area (IA) = track 4 IA + track 5 IA = track4&5 IA
track4buff <- inun_2010 %>% 
  crop(extent(trk_2010_buff[7,])) %>% 
  mask(trk_2010_buff[7,])
track5buff <- inun_2010 %>% 
  crop(extent(trk_2010_buff[4,])) %>% 
  mask(trk_2010_buff[4,])

plot(track4buff)
plot(track5buff)

area_2010_T4_inun <- as.data.frame(freq(track4buff)) %>% 
  subset(value == 1)
area_2010_T5_inun <- as.data.frame(freq(track5buff)) %>% 
  subset(value == 1)

overlapinun2010 <- (area_2010_T4_inun$count + 
    area_2010_T5_inun$count - 
    area_2010_T4_5_inun$count)*30*30/10000

# revised track 4 and 5 areas
track4area <- as.numeric((st_area(trk_2010_buff[7,])/10000)) - 0.5*overlaparea2010 # track 4
round(track4area)
track5area <- as.numeric((st_area(trk_2010_buff[4,])/10000)) - 0.5*overlaparea2010 # track 5
round(track5area)

# revised track 4 and 5 inundated areas
track4inun <- area_2010_T4_inun$count*30*30/10000 - 0.5*overlapinun2010 # track 4
track5inun <- area_2010_T5_inun$count*30*30/10000 - 0.5*overlapinun2010 # track 5
 
# revised track 4 and 5 inundated proportions
track4prop <- track4inun/track4area*100
track4prop
track5prop <- track5inun/track5area*100
track5prop




# check other years
plot(inun_2011)
plot(trk_2011_C1_all_buff, add = TRUE)
plot(trk_2011_blind_C1_buff, add = TRUE)

plot(inun_2012)
plot(trk_2012_C1_buff, add = TRUE)

plot(inun_2013)
plot(trk_2013_C1_buff, add = TRUE)

plot(inun_2014)
plot(trk_2014_C1_all_buff, add = TRUE)
plot(trk_2014_blind_C1_buff, add = TRUE)

plot(inun_2015)
plot(trk_2015_main_C1_buff, add = TRUE)
plot(trk_2015_Bac_C1_buff, add = TRUE)

plot(inun_2016)
plot(trk_2016_transect_buff, add = TRUE)
plot(trk_20161107_Bac_C1_buff, add = TRUE)
plot(trk_20161109_C1_buff, add = TRUE)

plot(inun_2017)
plot(trk_2017_C1_all_buff, add = TRUE)
plot(trk_2017_Bac_buff, add = TRUE)
plot(trk_2017_blind_C1_buff, add = TRUE)

plot(inun_2018)
plot(trk_2018_main_C1_buff, add = TRUE)
plot(trk_2018_Bac_buff, add = TRUE)

plot(inun_2019)
plot(trk_2019_buff, add = TRUE)

plot(inun_2020)
plot(trk_2020_main_C1_buff, add = TRUE)
plot(trk_2020_Bac_C1_all_buff, add = TRUE)

plot(inun_2021)
plot(trk_2021_main_C1_buff, add = TRUE)
plot(trk_2021_Bac_buff, add = TRUE)

plot(inun_2022)
plot(trk_2022_buff, add = TRUE)

plot(inun_2023)
plot(trk_2023_C1buff_all, add = TRUE)
plot(trk_2023_Bac_buff, add = TRUE)

plot(inun_2024)
plot(trk_2024_C1_buff, add = TRUE)
plot(trk_2024_C1t_buff, add = TRUE)
plot(trk_2024_Baconian_buff, add = TRUE)

```

```{r Overlap_areas_2016}

# 2016 - track 6 & 7 overlap

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2016_buff_T6_7_inun <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[c(9,10),])) %>% 
  mask(trk_20161109_C1_all[c(9,10),])

area_2016_T6_7 <- as.data.frame(freq(trk_2016_buff_T6_7_inun)) 
area_2016_T6_7_inun <- area_2016_T6_7 %>% subset(value == 1)
area_2016_T6_7_noninun <- area_2016_T6_7 %>% subset(value == 0)

# overlapped area = track 6 area + track 7 area - track6&7 area
plot(trk_2016_buff_T6_7_inun)

overlaparea2016_T6_7 <- as.numeric(st_area(trk_20161109_C1_all[9,])/10000) + 
  as.numeric(st_area(trk_20161109_C1_all[10,])/10000) - 
  (area_2016_T6_7_inun$count + area_2016_T6_7_noninun$count)*30*30/10000 # account for cell size

# overlapped inundation area (IA) = track 6 IA + track 7 IA = track6&7 IA
track6buff <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[9,])) %>% 
  mask(trk_20161109_C1_all[9,])
track7buff <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[10,])) %>% 
  mask(trk_20161109_C1_all[10,])

plot(track6buff)
plot(track7buff)

area_2016_T6_inun <- as.data.frame(freq(track6buff)) %>% 
  subset(value == 1)
area_2016_T7_inun <- as.data.frame(freq(track7buff)) %>% 
  subset(value == 1)

overlapinun2016 <- (area_2016_T6_inun$count + 
    area_2016_T7_inun$count - 
    area_2016_T6_7_inun$count)*30*30/10000

# revised track 6 and 7 areas
track6area <- as.numeric((st_area(trk_20161109_C1_all[9,])/10000)) - 0.5*overlaparea2016_T6_7 # track 6
round(track6area)
track7area <- as.numeric((st_area(trk_20161109_C1_all[10,])/10000)) - 0.5*overlaparea2016_T6_7 # track 7
round(track7area)

# revised track 6 and 7 inundated areas
track6inun <- area_2016_T6_inun$count*30*30/10000 - 0.5*overlapinun2016 # track 6
track7inun <- area_2016_T7_inun$count*30*30/10000 - 0.5*overlapinun2016 # track 7
 
# revised track 6 and 7 inundated proportions
track6prop <- track6inun/track6area*100
track6prop
track7prop <- track7inun/track7area*100
track7prop





# 2016 - 3a & track 8 and track 9

# prepare by creating a summary of inundated and non inundated areas in the overlaps
trk_2016_buff_T8_3A_inun <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[c(2,3),])) %>% 
  mask(trk_20161109_C1_all[c(2,3),])

area_2016_T8_3A <- as.data.frame(freq(trk_2016_buff_T8_3A_inun)) 
area_2016_T8_3A_inun <- area_2016_T8_3A %>% subset(value == 1)
area_2016_T8_3A_noninun <- area_2016_T8_3A %>% subset(value == 0)

trk_2016_buff_T9_3A_inun <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[c(2,11),])) %>% 
  mask(trk_20161109_C1_all[c(2,11),])

area_2016_T9_3A <- as.data.frame(freq(trk_2016_buff_T9_3A_inun)) 
area_2016_T9_3A_inun <- area_2016_T9_3A %>% subset(value == 1)
area_2016_T9_3A_noninun <- area_2016_T9_3A %>% subset(value == 0)

# overlapped area 1 = 3A area + track 8 area - track3A&8 area
plot(trk_2016_buff_T8_3A_inun)

overlaparea2016_T8_3A <- as.numeric(st_area(trk_20161109_C1_all[2,])/10000) + 
  as.numeric(st_area(trk_20161109_C1_all[3,])/10000) - 
  (area_2016_T8_3A_inun$count + area_2016_T8_3A_noninun$count)*30*30/10000 # account for cell size

# overlapped area 2 = 3A area + track 9 area - track3A&9 area
plot(trk_2016_buff_T9_3A_inun)

overlaparea2016_T9_3A <- as.numeric(st_area(trk_20161109_C1_all[2,])/10000) + 
  as.numeric(st_area(trk_20161109_C1_all[11,])/10000) - 
  (area_2016_T9_3A_inun$count + area_2016_T8_3A_noninun$count)*30*30/10000 # account for cell size


# overlapped inundation area (IA) 1 = 3A IA + track 8 IA = track3A&8 IA
track3Abuff <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[2,])) %>% 
  mask(trk_20161109_C1_all[2,])
track8buff <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[3,])) %>% 
  mask(trk_20161109_C1_all[3,])

plot(track3Abuff)
plot(track8buff)

area_2016_3A_inun <- as.data.frame(freq(track3Abuff)) %>% 
  subset(value == 1)
area_2016_T8_inun <- as.data.frame(freq(track8buff)) %>% 
  subset(value == 1)

overlapinun2016_3A_T8 <- (area_2016_3A_inun$count + 
    area_2016_T8_inun$count - 
    area_2016_T8_3A_inun$count)*30*30/10000

##
track9buff <- inun_2016 %>% 
  crop(extent(trk_20161109_C1_all[11,])) %>% 
  mask(trk_20161109_C1_all[11,])

plot(track9buff)

area_2016_T9_inun <- as.data.frame(freq(track9buff)) %>% 
  subset(value == 1)

overlapinun2016_3A_T9 <- (area_2016_3A_inun$count + 
    area_2016_T9_inun$count - 
    area_2016_T9_3A_inun$count)*30*30/10000


# revised 3A, track 8 and 9 areas
track3Aarea <- as.numeric((st_area(trk_20161109_C1_all[2,])/10000)) - 0.5*overlaparea2016_T8_3A - 0.5*overlaparea2016_T9_3A # 3A
round(track3Aarea)
track8area <- as.numeric((st_area(trk_20161109_C1_all[3,])/10000)) - 0.5*overlaparea2016_T8_3A # track 8
round(track8area)
track9area <- as.numeric((st_area(trk_20161109_C1_all[11,])/10000)) - 0.5*overlaparea2016_T9_3A # track 9
round(track9area)

# revised 3A, track 8 and 9inundated areas
track3Ainun <- area_2016_3A_inun$count*30*30/10000 - 0.5*overlapinun2016_3A_T8 - 0.5*overlapinun2016_3A_T9 # track 3A
track8inun <- area_2016_T8_inun$count*30*30/10000 - 0.5*overlapinun2016_3A_T8 # track 8
track9inun <- area_2016_T9_inun$count*30*30/10000 - 0.5*overlapinun2016_3A_T9 # track 9
 
# revised 3A, track 8 and 9inundated proportions
track3Aprop <- track3Ainun/track3Aarea*100
track3Aprop
track8prop <- track8inun/track8area*100
track8prop
track9prop <- track9inun/track9area*100
track9prop

```

```{r Overlap_areas_2017}
# first, visualise transects
plot(inun_2017)
plot(trk_2017_C1_all_buff, add = TRUE)
plot(trk_2017_Bac_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2017_buff_C3_LL_inun <- inun_2017 %>% 
  crop(extent(trk_2017_C1_all_buff[c(3,2),])) %>% 
  mask(trk_2017_C1_all_buff[c(3,2),])

area_2017_C3_LL <- as.data.frame(freq(trk_2017_buff_C3_LL_inun)) 
area_2017_C3_LL_inun <- area_2017_C3_LL %>% subset(value == 1)
area_2017_C3_LL_noninun <- area_2017_C3_LL %>% subset(value == 0)

# overlapped area = C3 area + LL area - trackC3&LL area
plot(trk_2017_buff_C3_LL_inun)

overlaparea2017 <- as.numeric(st_area(trk_2017_C1_all_buff[3,])/10000) + 
  as.numeric(st_area(trk_2017_C1_all_buff[2,])/10000) - 
  (area_2017_C3_LL_inun$count + area_2017_C3_LL_noninun$count)*30*30/10000 # account for cell size

# overlapped inundation area (IA) = track C3 IA + track LL IA = trackC3&LL IA
trackC3buff <- inun_2017 %>% 
  crop(extent(trk_2017_C1_all_buff[3,])) %>% 
  mask(trk_2017_C1_all_buff[3,])
trackLLbuff <- inun_2017 %>% 
  crop(extent(trk_2017_C1_all_buff[2,])) %>% 
  mask(trk_2017_C1_all_buff[2,])

plot(trackC3buff)
plot(trackLLbuff)

area_2017_C3_inun <- as.data.frame(freq(trackC3buff)) %>% 
  subset(value == 1)
area_2017_LL_inun <- as.data.frame(freq(trackLLbuff)) %>% 
  subset(value == 1)

overlapinun2017 <- (area_2017_LL_inun$count + 
    area_2017_C3_inun$count - 
    area_2017_C3_LL_inun$count)*30*30/10000

# revised C3 and LL areas
trackC3area <- as.numeric((st_area(trk_2017_C1_all_buff[3,])/10000)) - 0.5*overlaparea2017 #  Cumbung 3
round(trackC3area)
trackLLarea <- as.numeric((st_area(trk_2017_C1_all_buff[2,])/10000)) - 0.5*overlaparea2017 # Lignum Lake Not
round(trackLLarea)

# revised C3 and LL inundated areas
trackC3inun <- area_2017_C3_inun$count*30*30/10000 - 0.5*overlapinun2017 # track 4
trackLLinun <- area_2017_LL_inun$count*30*30/10000 - 0.5*overlapinun2017 # track 5
 
# revised C3 and LL inundated proportions
trackC3prop <- trackC3inun/trackC3area*100
trackC3prop
trackLLprop <- trackLLinun/trackLLarea*100
trackLLprop

```

\

```{r Overlap_areas_2018}
# first, visualise transects
plot(inun_2018)
plot(trk_2018_main_C1_buff, add = TRUE)
plot(trk_2018_Bac_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2018_buff_GCS_GCS2_inun <- inun_2018 %>% 
  crop(extent(trk_2018_main_C1_buff[c(1,2),])) %>% 
  mask(trk_2018_main_C1_buff[c(1,2),])

area_2018_GCS_GCS2 <- as.data.frame(freq(trk_2018_buff_GCS_GCS2_inun)) 
area_2018_GCS_GCS2_inun <- area_2018_GCS_GCS2 %>% subset(value == 1)
area_2018_GCS_GCS2_noninun <- area_2018_GCS_GCS2 %>% subset(value == 0)

# overlapped area = GCS area + GCS2 area - GCS&GCS2 area
plot(trk_2018_buff_GCS_GCS2_inun)

overlaparea2018 <- as.numeric(st_area(trk_2018_main_C1_buff[1,])/10000) + 
  as.numeric(st_area(trk_2018_main_C1_buff[2,])/10000) - 
  (area_2018_GCS_GCS2_inun$count + area_2018_GCS_GCS2_noninun$count)*30*30/10000 # account for cell size

# overlapped inundation area (IA) = GCS IA + GCS2 IA - GCS&GCS2 IA
GCSbuff <- inun_2018 %>% 
  crop(extent(trk_2018_main_C1_buff[1,])) %>% 
  mask(trk_2018_main_C1_buff[1,])
GCS2buff <- inun_2018 %>% 
  crop(extent(trk_2018_main_C1_buff[2,])) %>% 
  mask(trk_2018_main_C1_buff[2,])

plot(GCSbuff)
plot(GCS2buff)

area_2018_GCS_inun <- as.data.frame(freq(GCSbuff)) %>% 
  subset(value == 1)
area_2018_GCS2_inun <- as.data.frame(freq(GCS2buff)) %>% 
  subset(value == 1)

overlapinun2018 <- (area_2018_GCS_inun$count + 
    area_2018_GCS2_inun$count - 
    area_2018_GCS_GCS2_inun$count)*30*30/10000

# revised GCS and GCS2 areas
GCSarea <- as.numeric((st_area(trk_2018_main_C1_buff[1,])/10000)) - 0.5*overlaparea2018 # GCS
round(GCSarea)
GCS2area <- as.numeric((st_area(trk_2018_main_C1_buff[2,])/10000)) - 0.5*overlaparea2018 # GCS2
round(GCS2area)

# revised GCS and GCS2  inundated areas
GCSinun <- area_2018_GCS_inun$count*30*30/10000 - 0.5*overlapinun2018 # track 4
GCS2inun <- area_2018_GCS2_inun$count*30*30/10000 - 0.5*overlapinun2018 # track 5
 
# revised GCS and GCS2  inundated proportions
GCSprop <- GCSinun/GCSarea*100
GCSprop
GCS2prop <- GCS2inun/GCS2area*100
GCS2prop


## Count 2

# first, visualise transects
plot(inun_2018)
plot(trk_2018_main_C2_buff, add = TRUE)
plot(trk_2018_Bac_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2018_buff_GCS_GCS2_inun <- inun_2018 %>% 
  crop(extent(trk_2018_main_C2_buff[c(1,3),])) %>% 
  mask(trk_2018_main_C2_buff[c(1,3),])

area_2018_GCS_GCS2 <- as.data.frame(freq(trk_2018_buff_GCS_GCS2_inun)) 
area_2018_GCS_GCS2_inun <- area_2018_GCS_GCS2 %>% subset(value == 1)
area_2018_GCS_GCS2_noninun <- area_2018_GCS_GCS2 %>% subset(value == 0)

# overlapped area = GCS area + GCS2 area - GCS&GCS2 area
plot(trk_2018_buff_GCS_GCS2_inun)

overlaparea2018 <- as.numeric(st_area(trk_2018_main_C2_buff[1,])/10000) + 
  as.numeric(st_area(trk_2018_main_C2_buff[3,])/10000) - 
  (area_2018_GCS_GCS2_inun$count + area_2018_GCS_GCS2_noninun$count)*30*30/10000 # account for cell size

# overlapped inundation area (IA) = GCS IA + GCS2 IA - GCS&GCS2 IA
GCSbuff <- inun_2018 %>% 
  crop(extent(trk_2018_main_C2_buff[1,])) %>% 
  mask(trk_2018_main_C2_buff[1,])
GCS2buff <- inun_2018 %>% 
  crop(extent(trk_2018_main_C2_buff[3,])) %>% 
  mask(trk_2018_main_C2_buff[3,])

plot(GCSbuff)
plot(GCS2buff)

area_2018_GCS_inun <- as.data.frame(freq(GCSbuff)) %>% 
  subset(value == 1)
area_2018_GCS2_inun <- as.data.frame(freq(GCS2buff)) %>% 
  subset(value == 1)

overlapinun2018 <- (area_2018_GCS_inun$count + 
    area_2018_GCS2_inun$count - 
    area_2018_GCS_GCS2_inun$count)*30*30/10000

# revised GCS and GCS2 areas
GCSarea <- as.numeric((st_area(trk_2018_main_C2_buff[1,])/10000)) - 0.5*overlaparea2018 # GCS
round(GCSarea)
GCS2area <- as.numeric((st_area(trk_2018_main_C2_buff[3,])/10000)) - 0.5*overlaparea2018 # GCS2
round(GCS2area)

# revised GCS and GCS2  inundated areas
GCSinun <- area_2018_GCS_inun$count*30*30/10000 - 0.5*overlapinun2018 # track 4
GCS2inun <- area_2018_GCS2_inun$count*30*30/10000 - 0.5*overlapinun2018 # track 5
 
# revised GCS and GCS2  inundated proportions
GCSprop <- GCSinun/GCSarea*100
GCSprop
GCS2prop <- GCS2inun/GCS2area*100
GCS2prop
```

```{r Overlap_2019}
# first, visualise transects
plot(inun_2019)
plot(trk_2019_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2019_buff_C3_GCS_inun <- inun_2019 %>% 
  crop(extent(trk_2019_buff[c(2,3),])) %>% 
  mask(trk_2019_buff[c(2,3),])

area_2019_C3_GCS <- as.data.frame(freq(trk_2019_buff_C3_GCS_inun)) 
area_2019_C3_GCS_inun <- area_2019_C3_GCS %>% subset(value == 1)
area_2019_C3_GCS_noninun <- area_2019_C3_GCS %>% subset(value == 0)

# overlapped area = Cumbung 3 area + GCS area - combined area
plot(trk_2019_buff_C3_GCS_inun)

overlaparea2019 <- as.numeric(st_area(trk_2019_buff[2,])/10000) + 
  as.numeric(st_area(trk_2019_buff[3,])/10000) - 
  (area_2019_C3_GCS_inun$count + area_2019_C3_GCS_noninun$count)*10*10/10000 # account for cell size (10*10m sentinel)

# overlapped inundation area (IA) = Cumbung 3 IA + GCS IA = combined IA
trackC3buff <- inun_2019 %>% 
  crop(extent(trk_2019_buff[2,])) %>% 
  mask(trk_2019_buff[2,])
trackGCSbuff <- inun_2019 %>% 
  crop(extent(trk_2019_buff[3,])) %>% 
  mask(trk_2019_buff[3,])

plot(trackC3buff)
plot(trackGCSbuff)

area_2019_C3_inun <- as.data.frame(freq(trackC3buff)) %>% 
  subset(value == 1)
area_2019_GCS_inun <- as.data.frame(freq(trackGCSbuff)) %>% 
  subset(value == 1)

overlapinun2019 <- (area_2019_C3_inun$count + 
    area_2019_GCS_inun$count - 
    area_2019_C3_GCS_inun$count)*10*10/10000

# revised track C3 and GCS areas
trackC3area <- as.numeric((st_area(trk_2019_buff[2,])/10000)) - 0.5*overlaparea2019 # track 4
round(trackC3area)
trackGCSarea <- as.numeric((st_area(trk_2019_buff[3,])/10000)) - 0.5*overlaparea2019 # track 5
round(trackGCSarea)

# revised track C3 and GCS inundated areas
trackC3inun <- area_2019_C3_inun$count*10*10/10000 - 0.5*overlapinun2019 # trackC3
trackGCSinun <- area_2019_GCS_inun$count*10*10/10000 - 0.5*overlapinun2019 # track GCS
 
# revised track C3 and GCS inundated proportions
trackC3prop <- trackC3inun/trackC3area*100
trackC3prop
trackGCSprop <- trackGCSinun/trackGCSarea*100
trackGCSprop

```

```{r Overlap_2020}
# first, visualise transects
plot(inun_2020)
plot(trk_2020_main_C1_buff, add = TRUE)
plot(trk_2020_Bac_C1_all_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2020_buff_E1_C3_inun <- inun_2020 %>% 
  crop(extent(trk_2020_main_C1_buff[c(1,5),])) %>% 
  mask(trk_2020_main_C1_buff[c(1,5),])

area_2020_E1_C3 <- as.data.frame(freq(trk_2020_buff_E1_C3_inun)) 
area_2020_E1_C3_inun <- area_2020_E1_C3 %>% subset(value == 1)
area_2020_E1_C3_noninun <- area_2020_E1_C3 %>% subset(value == 0)

# overlapped area = E1 area + C3 area - combined area
plot(trk_2020_buff_E1_C3_inun)

overlaparea2020 <- as.numeric(st_area(trk_2020_main_C1_buff[1,])/10000) + 
  as.numeric(st_area(trk_2020_main_C1_buff[5,])/10000) - 
  (area_2020_E1_C3_inun$count + area_2020_E1_C3_noninun$count)*10*10/10000 # account for cell size

# overlapped inundation area (IA) = E1 IA + C3 IA = combined IA
trackE1buff <- inun_2020 %>% 
  crop(extent(trk_2020_main_C1_buff[1,])) %>% 
  mask(trk_2020_main_C1_buff[1,])
trackC3buff <- inun_2020 %>% 
  crop(extent(trk_2020_main_C1_buff[5,])) %>% 
  mask(trk_2020_main_C1_buff[5,])

plot(trackE1buff)
plot(trackC3buff)

area_2020_E1_inun <- as.data.frame(freq(trackE1buff)) %>% 
  subset(value == 1)
area_2020_C3_inun <- as.data.frame(freq(trackC3buff)) %>% 
  subset(value == 1)

overlapinun2020 <- (area_2020_E1_inun$count + 
    area_2020_C3_inun$count - 
    area_2020_E1_C3_inun$count)*10*10/10000

# revised E1 & C3 areas
trackE1area <- as.numeric((st_area(trk_2020_main_C1_buff[1,])/10000)) - 0.5*overlaparea2020 # track E1
round(trackE1area)
trackC3area <- as.numeric((st_area(trk_2020_main_C1_buff[5,])/10000)) - 0.5*overlaparea2020 # track 5
round(trackC3area)

# revised E1 & C3 inundated areas
trackE1inun <- area_2020_E1_inun$count*10*10/10000 - 0.5*overlapinun2020 # E1
trackC3inun <- area_2020_C3_inun$count*10*10/10000 - 0.5*overlapinun2020 # C3
 
# revised E1 & C3 inundated proportions
trackE1prop <- trackE1inun/trackE1area*100
trackE1prop
trackC3prop <- trackC3inun/trackC3area*100
trackC3prop


# Count 2

# first, visualise transects
plot(inun_2020)
plot(trk_2020_main_C2_buff, add = TRUE)
plot(trk_2020_Bac_C2_all_buff, add = TRUE)

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2020_buff_E1_C3_inun <- inun_2020 %>% 
  crop(extent(trk_2020_main_C2_buff[c(1,4),])) %>% 
  mask(trk_2020_main_C2_buff[c(1,4),])

area_2020_E1_C3 <- as.data.frame(freq(trk_2020_buff_E1_C3_inun)) 
area_2020_E1_C3_inun <- area_2020_E1_C3 %>% subset(value == 1)
area_2020_E1_C3_noninun <- area_2020_E1_C3 %>% subset(value == 0)

# overlapped area = E1 area + C3 area - combined area
plot(trk_2020_buff_E1_C3_inun)

overlaparea2020 <- as.numeric(st_area(trk_2020_main_C2_buff[1,])/10000) + 
  as.numeric(st_area(trk_2020_main_C2_buff[4,])/10000) - 
  (area_2020_E1_C3_inun$count + area_2020_E1_C3_noninun$count)*10*10/10000 # account for cell size

# overlapped inundation area (IA) = E1 IA + C3 IA = combined IA
trackE1buff <- inun_2020 %>% 
  crop(extent(trk_2020_main_C2_buff[1,])) %>% 
  mask(trk_2020_main_C2_buff[1,])
trackC3buff <- inun_2020 %>% 
  crop(extent(trk_2020_main_C2_buff[4,])) %>% 
  mask(trk_2020_main_C2_buff[4,])

plot(trackE1buff)
plot(trackC3buff)

area_2020_E1_inun <- as.data.frame(freq(trackE1buff)) %>% 
  subset(value == 1)
area_2020_C3_inun <- as.data.frame(freq(trackC3buff)) %>% 
  subset(value == 1)

overlapinun2020 <- (area_2020_E1_inun$count + 
    area_2020_C3_inun$count - 
    area_2020_E1_C3_inun$count)*10*10/10000

# revised E1 & C3 areas
trackE1area <- as.numeric((st_area(trk_2020_main_C2_buff[1,])/10000)) - 0.5*overlaparea2020 # track E1
round(trackE1area)
trackC3area <- as.numeric((st_area(trk_2020_main_C2_buff[4,])/10000)) - 0.5*overlaparea2020 # track 5
round(trackC3area)

# revised E1 & C3 inundated areas
trackE1inun <- area_2020_E1_inun$count*10*10/10000 - 0.5*overlapinun2020 # E1
trackC3inun <- area_2020_C3_inun$count*10*10/10000 - 0.5*overlapinun2020 # C3
 
# revised E1 & C3 inundated proportions
trackE1prop <- trackE1inun/trackE1area*100
trackE1prop
trackC3prop <- trackC3inun/trackC3area*100
trackC3prop
```

2025\.

```{r overlap_2025}
# first, visualise transects
plot(inun_2025)
plot(trk_2025_buff, add = TRUE)

# transect 3 & 5 overlap

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2025_buff_3_5_inun <- inun_2025 %>% 
  crop(extent(trk_2025_buff[c(3,5),])) %>% 
  mask(trk_2025_buff[c(3,5),])

area_2025_3_5_ <- as.data.frame(freq(trk_2025_buff_3_5_inun)) 
area_2025_3_5_inun <- area_2025_3_5_ %>% subset(value == 1)
area_2025_3_5_noninun <- area_2025_3_5_ %>% subset(value == 0)

# overlapped area = transect 3 area + transect 5 area - combined area
plot(trk_2025_buff_3_5_inun)

overlaparea2025 <- as.numeric(st_area(trk_2025_buff[3,])/10000) + 
  as.numeric(st_area(trk_2025_buff[5,])/10000) - 
  (area_2025_3_5_inun$count + area_2025_3_5_noninun$count)*30*30/10000 # account for cell size (30 * 30 as Landsat)

# overlapped inundation area (IA) = T3 IA + T5 IA = combined IA
track3buff <- inun_2025 %>% 
  crop(extent(trk_2025_buff[3,])) %>% 
  mask(trk_2025_buff[3,])
track5buff <- inun_2025 %>% 
  crop(extent(trk_2025_buff[5,])) %>% 
  mask(trk_2025_buff[5,])

plot(track3buff)
plot(track5buff)

area_2025_T3_inun <- as.data.frame(freq(track3buff)) %>% 
  subset(value == 1)
area_2025_T5_inun <- as.data.frame(freq(track5buff)) %>% 
  subset(value == 1)

overlapinun2025 <- (area_2025_T3_inun$count + 
    area_2025_T5_inun$count - 
    area_2025_3_5_inun$count)*30*30/10000

# revised transect 3 & 5  areas
track3area <- as.numeric((st_area(trk_2025_buff[3,])/10000)) - 0.5*overlaparea2025 # track E1
round(track3area)
track5area <- as.numeric((st_area(trk_2025_buff[5,])/10000)) - 0.5*overlaparea2025 # track 5
round(track5area)

# revised transect 3 & 5  inundated areas
track3inun <- area_2025_T3_inun$count*30*30/10000 - 0.5*overlapinun2025 # E1
track5inun <- area_2025_T5_inun$count*30*30/10000 - 0.5*overlapinun2025 # C3
 
# revised transect 3 & 5 inundated proportions
track3prop <- track3inun/track3area*100
track3prop
track5prop <- track5inun/track5area*100
track5prop



# transects 4 & 6 overlap

# prepare by creating a summary of inundated and non inundated areas in the overlap
trk_2025_buff_4_6_inun <- inun_2025 %>% 
  crop(extent(trk_2025_buff[c(4,6),])) %>% 
  mask(trk_2025_buff[c(4,6),])

area_2025_4_6_ <- as.data.frame(freq(trk_2025_buff_4_6_inun)) 
area_2025_4_6_inun <- area_2025_4_6_ %>% subset(value == 1)
area_2025_4_6_noninun <- area_2025_4_6_ %>% subset(value == 0)

# overlapped area = transect 4 area + transect 6 area - combined area
plot(trk_2025_buff_4_6_inun)

overlaparea2025 <- as.numeric(st_area(trk_2025_buff[4,])/10000) + 
  as.numeric(st_area(trk_2025_buff[6,])/10000) - 
  (area_2025_4_6_inun$count + area_2025_4_6_noninun$count)*30*30/10000 # account for cell size (30 * 30 as Landsat)

# overlapped inundation area (IA) = T4 IA + T6 IA = combined IA
track4buff <- inun_2025 %>% 
  crop(extent(trk_2025_buff[4,])) %>% 
  mask(trk_2025_buff[4,])
track6buff <- inun_2025 %>% 
  crop(extent(trk_2025_buff[6,])) %>% 
  mask(trk_2025_buff[6,])

plot(track4buff)
plot(track6buff)

area_2025_T4_inun <- as.data.frame(freq(track4buff)) %>% 
  subset(value == 1)
area_2025_T6_inun <- as.data.frame(freq(track6buff)) %>% 
  subset(value == 1)

overlapinun2025 <- (area_2025_T4_inun$count + 
    area_2025_T6_inun$count - 
    area_2025_4_6_inun$count)*30*30/10000

# revised transect 4 & 6 areas
track4area <- as.numeric((st_area(trk_2025_buff[4,])/10000)) - 0.5*overlaparea2025 # track 4
round(track4area)
track6area <- as.numeric((st_area(trk_2025_buff[6,])/10000)) - 0.5*overlaparea2025 # track 6
round(track6area)

# revised transect 4 & 6 inundated areas
track4inun <- area_2025_T4_inun$count*30*30/10000 - 0.5*overlapinun2025 # 4
track6inun <- area_2025_T6_inun$count*30*30/10000 - 0.5*overlapinun2025 # 6
 
# revised E1 & C3 inundated proportions
track4prop <- track4inun/track4area*100
track4prop
track6prop <- track6inun/track6area*100
track6prop

```

## 5. Waterbird calculations

Use the proportion inundated to calculate waterbird numbers for each species for the cumbung ecological boundary. Also keep current counts for lakes. Put these all in a table with total counts for each species.

These are examples - ongoing, we are doing the waterbird count part in The Duck Web.

```{r Waterbird_calculations}
wb_2010_sub <- wb_2010 %>% 
  subset(Date == "2010-11-17") %>% 
  subset(NewSubWetland == "Cumbung Swamp") %>% 
  subset(ReplicateCount == 1) 

wb_2010_summary <- wb_2010_sub %>% 
  group_by(SPP) %>% 
  summarise(sum = sum(Count)) %>% 
  mutate(extrap_count = round(sum * 100/19.35)) # 19.35 is the count 1 proportion of inundated areas surveyed for 2010

sum(wb_2010_sub$Count)
sum(wb_2010_summary$extrap_count)


wb_2011_sub <- wb_2011 %>% 
  subset(NewSubWetland %in% c("Cumbung Swamp", "Cumbung Swamp (Blindbungi Lake)")) %>%  
  subset(ReplicateCount == 1) 

wb_2011_summary <- wb_2011_sub %>% 
  group_by(SPP) %>% 
  summarise(sum = sum(Count)) %>% 
  mutate(extrap_count = round(sum * 100/35.8)) #35.8 is the count 1 proportion of inundated areas surveyed for 2011

sum(wb_2011_sub$Count)
sum(wb_2011_summary$extrap_count)


wb_2016_sub <- wb_2016 %>% 
  subset(Date == "2016-11-09") %>% 
  subset(NewSubWetland == "Cumbung Swamp") %>% 
  subset(ReplicateCount == 1) 

wb_2016_summary <- wb_2016_sub %>% 
  group_by(SPP) %>% 
  summarise(sum = sum(Count)) %>% 
  mutate(extrap_count = round(sum * 100/11.17)) # 11.17 is the count 1 proportion of inundated areas surveyed on 9 Nov 2016

sum(wb_2016_sub$Count)
sum(wb_2016_summary$extrap_count)
```
